---
title: "What If Analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{what_if_analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This brief introduction shows how use CMA to price alternative scenarios for an investment strategy. The recipe goes as follows:

1.  Detach the copulas from the margins;

2.  Pick one element to play with;

    -   Example: if the "Market" is normally distributed, the t-distribution could be used to see how VaR and CVaR changes;

3.  Generate new scenarios;

4.  Merge the newly generated scenarios into the empirical copulas.

For this vignette, the `EuStockMarket` dataset is used as *proxy* for the "Market".

The setup goes as follows:

```{r, message=FALSE, warning=FALSE}
# load packages
library(cma)     # core package
library(dplyr)   # data manipulation
library(tidyr)   # data manipulation
library(purrr)   # vetorized loops
library(ggplot2) # plotting

# compute log returns (stationarity)
x <- matrix(diff(log(EuStockMarkets)), ncol = 4)
n_row <- nrow(x)
head(x)
```

## Step 1. Separate the Copulas from the Margins

The first step evolves separating the margins from the copulas with `cma_separation()`.

```{r}
# First CMA Step
step_one <- cma_separation(x)
step_one
```

The output is a list with three components: the margins, the CDF, and the copula.

In this vignette, the margins are manipulated while keeping the copula structure unchanged.

## Step 2. Choose a Target Distribution

Once the the "separation" is done it's necessary to choose a target distribution for the data. For simplicity, let's fit a Student-t, but it could be *any* distribution.

```{r}
dist_t <- fit_t(step_one$marginal)
dist_t
```

## Step 3. Create New Scenarios

The next step requires some `tidyverse` skills.

Start building a list-column for the new scenarios wit `generate_margins()`:

```{r}
simul_tbl <- tibble::tibble(simulations = 1:100) |> 
    dplyr::mutate(new_scenarios = purrr::map(
        .x = rep(n_row, 100), 
        .f = ~ generate_margins(model = dist_t, n = .x)
        )
    ) |> 
    tidyr::unnest(cols = new_scenarios)
simul_tbl
```

The `simul_tbl` is a *tidy* (long-format) `tibble` with 100 marginal distributions that are consistent with the `dist_t` object fitted in step 2 (every row in `new_scenarios` holds a different simulation).

## Step 4. "Glue" the New Scenarios Into the Empirical Copula

To complete the scenario generation process it's important to "glue" the simulations back into the empirical copula that is stored in the `step_one` object.

```{r}
step_two <- simul_tbl |> 
    dplyr::mutate(cma_comb = purrr::map(
        .x = new_scenarios, 
        .f = ~ cma_combination(
            x      = .x,
            cdf    = step_one$cdf, 
            copula = step_one$copula)
        )
    )
step_two
```

The `cma_comb` column now holds 100 simulations that share a bond to each other, because their copula is the same.

Finally, it's possible to compute the performance of those portfolios over an equal-weight strategy, for example. It could be *any* strategy.

```{r}
step_two <- step_two |> 
    dplyr::mutate(weights = list(rep(0.25, 4)), 
                  pnl     = purrr::map2(.x = cma_comb, 
                                        .y = weights, 
                                        .f = ~ as.matrix(.x) %*% .y)) |> 
    dplyr::select(-c(cma_comb, weights))
step_two
```

Translate compounded returns to prices and plot:

```{r, fig.width=7, fig.height=4, fig.align='center'}
step_two |> 
    dplyr::mutate(pnl_prices = purrr::map(.x = pnl, .f = ~ cumprod(exp(.x))), 
                  pnl_ref    = list(cumprod(exp(x %*% rep(0.25, 4))))) |> 
    tidyr::unnest(cols = c(pnl_prices, pnl_ref)) |> 
    
    dplyr::group_by(simulations) |> 
    dplyr::mutate(rowid = 1:n_row) |> 
    dplyr::ungroup() |> 
    
    ggplot2::ggplot(ggplot2::aes(x = rowid, y = pnl_prices, group = simulations)) + 
    ggplot2::geom_line(col = "grey") + 
    ggplot2::geom_line(ggplot2::aes(y = pnl_ref), color = "purple") + 
    ggplot2::scale_y_log10(labels = scales::dollar_format()) + 
    ggplot2::labs(title    = "EuStockMarkets Equal Weight Strategy", 
                  subtitle = "Realized vs. Alternative Scenarios", 
                  x        = "Time Evolution", 
                  y        = "P&L")
```

Remember that all the paths above are compatible with the same marginal distribution specification. The difference between the "good" portfolios and the "bad" ones is purely due by chance.

Also note how the copula "ties" the ups and downs, that happens always at the same time, but with different magnitudes.
